% =============================================================== 
% Translation of book: shi_gong for diet                       *
% ===============================================================

\input epsf

%\documentstyle[doublespace,11pt]{report}
\documentstyle[11pt]{report}
\textheight 9.0in
\textwidth 6.30in
\voffset -0.85in
\hoffset -0.60in
\def\R{{\bf R}}
\def\j{{\rm j}}
\def\C{{\bf C}}
%\renewcommand{\thefigure}{\thechapter.\arabic{figure}}
\begin{document}

\title{\bf Chinese Shi-Gong For Diet }

\author{\bf S. Y. Zhou, S. Han, and K. Xu \\
        Translated by X. G. Chen}

\maketitle

\begin{abstract}
This book introduces a easy, effective method for diet -- shi-gong \footnote{ translator' note: 
Shi-Gong is a Chiniese word, shi means air, gong means gong-fu or martial art. Shi-Gong is a kind
of chinese martial art that can control the flow of air inside one's body.} method.
This method has no side effect,
and requires no medicine or tea, not any equipment. But it has fast and significant effect for diet.

In this book, the method and theory behind the method are introduced. It lists the important details
needed in practing shi-gong, practical cases study. It also includes the statistics, observation and
analysis of a diet class organized by EAST BEAUTY RESEARCH CENTER. And recipes for the food during
diet period are also provided by the book.  
\end{abstract}

\tableofcontents

\chapter{Preface}

"Fat" have become a common "illness" of many people, it makes young people loss slenderness,
it also easy to cause older ones to have high blood pressure, heart attack, back and leg pain and
many other illness. Therefor there is a wave of dieting around the world. All kinds of diet medicine,
or teatments appeared. These medicines or treatments can only work during a very short period of time. Once
you stoping these medicines or treatments, the body goes back to the normal weight. Even more, people afraid
that the medicines and treatments may cause side-effects and expect that a long, significant effective and no
side effect method could merge. The method introduced by this book is just such kind of method -- Shi-Gong Diet
method. It is a new development of oriental shi-gong and its another contribution to the mankind by the
ancient chinese shi-gong.  
\vspace{0.1 in}

The one of biggest feature of this method is that it is easy to learn: it may takes only couple of hours to learn it.
The method is not limited by the place or enviornment. It only very short period of time for each practice, and
every body can do it. The method has fast and great effect. It may take only 1-2 days to see the results, and
1-2 weeks to see significant result. One will have great benefit if he continues to practice. There are several
examples:
\begin{itemize}
\item Famous scientist Prof. Zhang, W., Vice Director of Chinese Science and Technology Committee,
former Vice President of Qihua University, lost 4 kg in 4 days (now lost 9 kg); 
\item Ms. Li, J., Famous anchor woman of Chinese Central Television (CCTV), lost 7 kg in 14 days (now lost 10 kg);
\item Mr. Mu, S., Vice Director of news department of Chinese Central Televesion (CCTV), lost 12.5 kg in 38 days ;
\item Prof. Fi, X. T., (hight 159 cm, weight 95 kg), Vice President of Chinese Peoples' Congress,
President of Chinese Democratic Association, said: "For me, solving the my fat balley problem is more difficult than
solving my academic problems". After one week's practice of this method, he surprisely found that his balley is
one hole smaller measured by his belt. Not long, he lost 6 kgs, that made him very happy.  
\end{itemize}

Now this easy, like magic, no side effect, shi-gong method has become popular inside the campus of Qihua university.
In a few months, CCTV, People's Daily Oversea's edition, China Daily, Gongming Daily, Senior Citizen Daily, River-City
Daily, China News (Taiwain), Wenhui Daily (Hong Kong), New Evening News (Hong Kong), Gonghua Daily (Malysia) and many
other newspaper and news agencies had reported the story about this method. Many readers sent letters expressing 
interests in the method. Some people from overseas are preparing to come to China to try the method. A new age that
human can control his body shape is coming not far away. This shi-gong diet method was invented by Zhou, S. Y.,
Han, S. and Xu, K. of Qihua University. It was originated from {\bf\em Frog breath Gong} of the Shi-Gong Master
Zhang, Y. L. of Jianxi Province. The book was written by Zhou and Han.  

\vspace {0.15 in}
The editor/author aknowledge the work by Mr. Wang, Q., P. in organizing and overall editing.

\vspace{0.25 in} 

Editors\\
1988 \\
Qihua University Campus

\chapter{Shi-Gong Diet Method: Method and Theory}
This shi-gong diet method consists three "Gong"s -- {\bf\em Frog breath gong (frog gong), Lotus seed sitting gong},
and {\bf\em frog wave gong}. Practicing frog wave gong can ease the hungry feeling that many people have during
practing. Practicing this shi-gong diet method, people can naturally eat less or stop eating during practicing period,
or eat small amount vegetables or fruits. Usually, if you do not eat food, it can produce the many effects such as
feeling weak, dazzle. But practicing this shi-gong, these effects can be reduced or eliminated, with no feeling of
hungary, and no side effect even you eat less or stop eating. This means that this method is not simply dieting, 
instead, it transfers the nutrition from one place to another through shi-gong. Therefore people who practing this
shi-gong, although eat less or not eat, can obtain the energy from the extra fat of the body. This is the mechanism
that this shi-gong can quickly help one lost weight and produces no side effects.
\vspace{0.15 in}

Practing sh-gong diet method to loss weight is different from simply dieting both in the mechanism and in the format.
Shi-Gong diet method only controls eating habit in a short period, while simple dieting has to control eating habit
over long period of time. The former lost fat very fast, while the latter slow. Besides first 3-5 days effect which
is different from before, the former will produce more energy for him than he had before he practices and no side
effect.  And the latter will harm one's health and the weight will recover one he stop dieting.

\section{Frog breath gong (Frog Gong)}
\subsubsection{Method}
Seting on bench or chair. The bench or chair is about one feet or 1 ft 2 in \footnote{Translator's note:
feet and inch used here are not U.S. length measures, They are Chinese length measures.
One Chinese feet = 1/3 meter, and one Chinese feet = 10 inches}, depending on the hight of one person, so that
there is about 90 degree between one's thigh, and lower limb. The foot is on the ground. The both knees are in
the position that the width between them is the same as the width of shoulder. For males. right hand as a fist,
left is hold outside of right hand (Female, left hand as a fist and right hand holds outside the left hand). Put
elbows on the top of knees, top of the body slightly inclines forward, lower head, put the eye of fist on the
forehead. Slightly close the eyes, and relax the body, make yourself comfortable.
\vspace{0.15 in}

After prepared above body positions, now begin to adjust one mind to make the thinking, neural systems into
a relax, quite status. First, sigh a breath, the body feels like so soft and weak, just a one has just experienced
a hard, fatigue labor. Then, try to think the most beautiful things in you life, make your spirit feels wonderful. 
Keeping soing this for 1-2 minutes, the body and mind will become status that natural, floating and wonderful. 

\vspace{0.15 in}

After finished the preparation for the mind, comes the main part of the shi-gong stage. Let the mind solely
concentrates on the activity of breathing, avoid the influence of any noise or things from outside -- 
even thunderstorm will not affects you~\footnote{Translator's note: This is a Chinese saying. Direct translation
is: even there is a thounder storm, you are not move. It usually used for analogy. It means that you are (or should)
so determined, or your are (should) so concentrated that even the thounder can not move you. Of course you should
move when thounder comes.}.
\vspace {0.05 in}

First take one mouth air into balley, then gradually take the air out. When send the
air out, one should slowly, with small stream of air, uniformly send it out. The body should relax while exhaling, 
and the air should comes out from lower portion of the balley, and one should feel the balley gradually
becomes soft.
\vspace {0.05 in}

After gradually exhaling, gradually inhales using the nose. Again, inhaling should be in a
small steam, uniform and slow. With the air takes in. The lower portion of the balley gradually expands, once
the lower balley expands to a certain extend, hold that position for 1-2 seconds, then have short exhale, then
quickly but gradually exhale the air. 
\vspace{0.05 in}

Repeat such process, ---- exhaling, inhaling, stopping 2 secconds, short exhaling, exhale, inhale, stoping for
2 second, short exhale, ... , --- continues as a no stoping cycles.  During the whole process, there is no
up and down moving in the cheast. Only the balley expands and extracts, just like a frog's balley expands and
extracts during breath, this the reason this shi-gong is called frog gong.

\vspace{0.05 in} 

\subsubsection{Theory}

 Add details later.

\section{Lutos seed sitting gong}
\section{Frog Wave gong}

\chapter{Frequent Asked Questions and Answers}
\chapter{Breif Introduction to Case Study}
\chapter{Shi-Gong Diet Class Report}
\chapter{Statistics, Observation and Analysis of Shi-Gong Diet Method}
\section{Introduction}
\section{Statistics}
\section{Observation and analysis}
\section{Discussion and Conclusions}
\chapter{Recipes of Food in Shi-Gong Diet Method}
\chapter{Appendix: Reports on Newspapers and Letters from Excisers}

\bigskip\bigskip 
%\bibliography{/homes/xgchen/SRC/bibliography/references} % specify bibliography in eosdis.bib
%\bibliographystyle{plain}                                % specify plain.sty as style file
%\bibliographystyle{unsrt}                                % specify plain.sty as style file
%\bibliographystyle{alpha}                                % specify plain.sty as style file
%\addcontentsline{toc}{chapter}{\protect\numberline{}{\bf REFERENCE}}

\end{document}
